---
title: "R in Action Regression Chapter 8"
output: html_notebook
---

Running the Model

```{r}
fit <- lm(formula, data)
```

formula
Y ~ X1 + X2 + ... + Xk

interaction between predictor variables
y ~ x + z + x:z

all possible interactions
y ~ x * z * w 
same as
y ~ x + z + w + x:z + x:w + z:w + x:z:w

interactions up to a specified degree
y ~ (x + z + w)^2
same as
y ~ x + z + w + x:z + x:w + z:w

all other variables in the data frame except the dependent variable
y ~ .
same as
y ~ x + z + w

removes a variable from the equation
y ~ (x + z + w)^2 – x:w
same as
y ~ x + z + w + x:z + z:w

Suppresses the intercept. forces the line through the origin at x=0
y ~ x -1

interpreted arithmetically
y ~ x + I((z + w)^2)
same as 
y ~ x + h
where h = (z + w)^2


```{r}
fit <- lm(weight ~ height, data=women)
summary(fit)

plot(women$height,women$weight,
     xlab="Height (in inches)",
     ylab="Weight (in pounds)")
abline(fit)

fit2 <- lm(weight ~ height + I(height^2), data=women)
summary(fit2)
plot(women$height,women$weight,
     xlab="Height (in inches)",
     ylab="Weight (in lbs)")
lines(women$height,fitted(fit2))

library(car)
scatterplot(weight ~ height, data=women,
            spread=FALSE, smoother.args=list(lty=2), pch=19,
            main="Women Age 30-39",
            xlab="Height (inches)",
            ylab="Weight (lbs.)")



states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
cor(states)
scatterplotMatrix(states, spread=FALSE, smoother.args=list(lty=2),
                  main="Scatter Plot Matrix")
fit <- lm(Murder ~ Population + Illiteracy + Income + Frost,
          data=states)
summary(fit)


fit <- lm(mpg ~ hp + wt + hp:wt, data=mtcars)
summary(fit)


#install.packages("effects")
library(effects)
plot(effect("hp:wt", fit,, list(wt=c(2.2,3.2,4.2))), multiline=TRUE)





states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)
confint(fit)


fit <- lm(weight ~ height, data=women)
par(mfrow=c(2,2))
plot(fit)


fit2 <- lm(weight ~ height + I(height^2), data=women)
par(mfrow=c(2,2))
plot(fit2)

newfit <- lm(weight~ height + I(height^2), data=women[-c(13,15),])
par(mfrow=c(2,2))
plot(newfit)


states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)
par(mfrow=c(2,2))
plot(fit)

```

# Diagnostics

## Checking Normality

```{r}
par(mfrow=c(1,1))

library(car)
states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
fit <- lm(Murder ~ Population + Illiteracy + Income + Frost, data=states)
qqPlot(fit, labels=row.names(states), id.method="identify",
       simulate=TRUE, main="Q-Q Plot")
states["Nevada",]
fitted(fit)["Nevada"]
residuals(fit)["Nevada"]

residplot <- function(fit, nbreaks=10) {
  z <- rstudent(fit)
  hist(z, breaks=nbreaks, freq=FALSE,
       xlab="Studentized Residual",
       main="Distribution of Errors")
  rug(jitter(z), col="brown")
  curve(dnorm(x, mean=mean(z), sd=sd(z)),
        add=TRUE, col="blue", lwd=2)
  lines(density(z)$x, density(z)$y,
        col="red", lwd=2, lty=2)
  legend("topright",
         legend = c( "Normal Curve", "Kernel Density Curve"),
         lty=1:2, col=c("blue","red"), cex=.7)
}
residplot(fit)

```


# Checking Independence

```{r}
durbinWatsonTest(fit)
```
# Checking Linearity

```{r}
crPlots(fit)
```


# Checking Variance
```{r}
crPlots(fit)

ncvTest(fit)
spreadLevelPlot(fit)
```


# Global validation of Linear Model
```{r}
#install.packages("gvlma")
library(gvlma)
gvmodel <- gvlma(fit)
summary(gvmodel)
```

# Checking Multicollinearity
```{r}
vif(fit)

sqrt(vif(fit)) > 2 # problem?
```

# Outliers
Unusually large residuals (Yi - Yi^)
```{r}
outlierTest(fit)
```

# High-leverage points
unusual vars
```{r}
hat.plot <- function(fit) {
  p <- length(coefficients(fit))
  n <- length(fitted(fit))
  plot(hatvalues(fit), main="Index Plot of Hat Values")
  abline(h=c(2,3)*p/n, col="red", lty=2)
  identify(1:n, hatvalues(fit), names(hatvalues(fit)))
}
hat.plot(fit)

cutoff <- 4/(nrow(states)-length(fit$coefficients)-2)
plot(fit, which=4, cook.levels=cutoff)
abline(h=cutoff, lty=2, col="red")

avPlots(fit, ask=FALSE)
```

# Checking if Transformation is needed
```{r}
influencePlot(fit, id.method="identify", main="Influence Plot",
              sub="Circle size is proportional to Cook's distance")


summary(powerTransform(states$Murder))

boxTidwell(Murder~Population+Illiteracy,data=states)
```


# Comparing Models
```{r}
states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
fit1 <- lm(Murder ~ Population + Illiteracy + Income + Frost,
           data=states)
fit2 <- lm(Murder ~ Population + Illiteracy, data=states)
anova(fit2, fit1)

AIC(fit1,fit2)
```


# stepwise regression
```{r}
library(MASS)
states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
fit <- lm(Murder ~ Population + Illiteracy + Income + Frost,
          data=states)
stepAIC(fit, direction="backward")
```

# Checking all combinations of vars
```{r}
#install.packages("leaps")
library(leaps)
states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
leaps <-regsubsets(Murder ~ Population + Illiteracy + Income +
                     Frost, data=states, nbest=4)
plot(leaps, scale="adjr2")
library(car)
subsets(leaps, statistic="cp",
        main="Cp Plot for All Subsets Regression")
abline(1,1,lty=2,col="red")
```


# Checking the consistency of predictions
```{r}
shrinkage <- function(fit, k=10){
  require(bootstrap)
  theta.fit <- function(x,y){lsfit(x,y)}
  theta.predict <- function(fit,x){cbind(1,x)%*%fit$coef}
  x <- fit$model[,2:ncol(fit$model)]
  y <- fit$model[,1]
  results <- crossval(x, y, theta.fit, theta.predict, ngroup=k)
  r2 <- cor(y, fit$fitted.values)^2
  r2cv <- cor(y, results$cv.fit)^2
  cat("Original R-square =", r2, "\n")
  cat(k, "Fold Cross-Validated R-square =", r2cv, "\n")
  cat("Change =", r2-r2cv, "\n")
}
```


# Checking the change of R-squared by building model on random selections of the data
```{r}
#install.packages("bootstrap")
library(bootstrap)
states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
fit <- lm(Murder ~ Population + Income + Illiteracy + Frost, data=states)
shrinkage(fit)
```

# Ranking the vars

```{r}
states <- as.data.frame(state.x77[,c("Murder", "Population",
                                     "Illiteracy", "Income", "Frost")])
zstates <- as.data.frame(scale(states))
zfit <- lm(Murder~Population + Income + Illiteracy + Frost, data=zstates)
coef(zfit)



relweights <- function(fit,...){
  R <- cor(fit$model)
  nvar <- ncol(R)
  rxx <- R[2:nvar, 2:nvar]
  rxy <- R[2:nvar, 1]
  svd <- eigen(rxx)
  evec <- svd$vectors
  ev <- svd$values
  delta <- diag(sqrt(ev))
  lambda <- evec %*% delta %*% t(evec)
  lambdasq <- lambda ^ 2
  beta <- solve(lambda) %*% rxy
  rsquare <- colSums(beta ^ 2)
  rawwgt <- lambdasq %*% beta ^ 2
  import <- (rawwgt / rsquare) * 100
  import <- as.data.frame(import)
  row.names(import) <- names(fit$model[2:nvar])
  names(import) <- "Weights"
  import <- import[order(import),1, drop=FALSE]
  dotchart(import$Weights, labels=row.names(import),
           xlab="% of R-Square", pch=19,
           main="Relative Importance of Predictor Variables",
           sub=paste("Total R-Square=", round(rsquare, digits=3)),
           ...)
  return(import)
}

relweights(fit)
```







